\begin{abstract}
Living organisms maintain complex shapes against gravity, but the physical limits of this maintenance are unknown. Here we derive a dimensionless scaling parameter --- the Bio-Gravitational Number $\mathcal{B}_g = EI/(MgL^2)$ --- that predicts when passive structure becomes insufficient and active metabolic maintenance is required. Cross-species analysis confirms that humans uniquely occupy an ``Allometric Trap'' ($\mathcal{B}_g \approx 0.01$) where shape maintenance depends entirely on metabolic energy. Using an Information--Elasticity Coupling framework that connects HOX-patterned developmental fields to Cosserat rod mechanics, we show that the spinal S-curve emerges as the energetic ground state, reproducing clinical lordosis (${\sim}42^\circ$) and kyphosis (${\sim}35^\circ$). AlphaFold structural analysis of 23 mechanotransduction proteins reveals a 72\% demand--supply anisotropy gap ($p = 0.011$): the proteins required for gravity sensing are thermodynamically expensive, while their metabolic regulators are intrinsically disordered and fragile. This molecular asymmetry creates an Energy Deficit Window during adolescent growth ($L > 0.35$\,m) that our model predicts produces scoliosis-like instabilities. These results reclassify Adolescent Idiopathic Scoliosis as a thermodynamic scaling failure and identify molecular targets for metabolic intervention.
The human spine maintains a complex S-shaped sagittal profile against gravity, yet its physical principle and adolescent vulnerability remain unexplained. We introduce the Information--Elasticity Coupling (IEC) framework, integrating a developmental information field representing HOX patterning with Cosserat rod mechanics, implemented via PyElastica simulations. The IEC framework selects the S-shaped spinal profile as the energetic ground state, reproducing normal lordosis and kyphosis. Phase diagram analysis reveals gravity-dominated, cooperative, and information-dominated morphological regimes. During simulated adolescent growth, small information-field asymmetries are amplified into scoliotic deformities through a supercritical bifurcation at a critical length, explaining adolescent scoliosis onset. Furthermore, AlphaFold structural analysis of mechanotransduction proteins reveals a demand-supply anisotropy gap quantifying this structural vulnerability. We conclude that spinal geometry emerges as a thermodynamic standing wave maintained by metabolic expenditure. The IEC framework unifies developmental patterning and biomechanics, generating specific testable predictions for scoliosis prevention.
\end{abstract}
