\section{Conclusions}

We have presented a theoretical framework for \emph{Biological Countercurvature}, positing that the geometry of the spine is determined by the interaction between a developmental information field and the gravitational environment. By coupling a Cosserat rod model with an Information--Elasticity mechanism, we demonstrated that the spinal S-curve emerges as a gravity-selected mode---a stable equilibrium accessible only when information modifies the effective metric of the structure.

This framework resolves five longstanding questions in spinal biology. First, the adolescent growth spurt is reinterpreted not as a design flaw that creates vulnerability, but as a gravitationally selected strategy to minimize total time in the Energy Deficit Window---organisms that transit rapidly accumulate less proprioceptive damage. Second, the diversity of AIS curve patterns (Lenke Types 1--6) is explained as distinct eigenmodes of the coupled Cosserat system, with mode selection determined by the regional distribution of protein expression gradients and the local energy deficit ratio. Third, the $\sim$10:1 female preponderance arises from metabolic timing rather than structural weakness: girls exhibit a deeper energy deficit ($R_{\mathrm{peak}} \approx 2.7$ vs.~2.4 in boys) due to unfavorable body composition and lower mitochondrial biogenesis capacity (PPARGC1A). Fourth, AlphaFold structural analysis of 23 proteins reveals a quantitative 34\% Demand--Supply Anisotropy Gap, a VIM Vulnerability Index of 3.01$\times$, and a Supply-Side Disorder Paradox (40\% vs.~28\%) that collectively explain the molecular cascade from energy depletion to scoliotic buckling. Fifth, the energy supply failure extends far beyond simple hunger, involving six distinct molecular mechanisms: a mitochondrial capacity ceiling, vascular supply limitation, circadian desynchrony, the modern height mismatch, micronutrient deficits, and a recursive supply constraint in which the supply machinery itself is expensive to maintain.

Together, these results unify the understanding of normal spinal development, microgravity adaptation, and pathological deformity under a single physical principle: the shaping of biological geometry by the interaction of genetic information and gravitational force. The framework yields falsifiable predictions spanning molecular biology (paraspinal PPARGC1A expression), clinical biomechanics (eigenmodes predict curve type from pre-radiographic muscle asymmetry), and therapeutic intervention (NAD$^+$ precursor supplementation, low-intensity vibration, circadian rhythm restoration). Ongoing work coupling the IEC framework to longitudinal gene expression atlases and prospective clinical cohorts will further test these predictions and pave the way for information-guided spinal medicine.
