\section{Methods}

We implement the Information--Cosserat framework using two complementary numerical approaches: a fast deterministic beam model for parameter sweeps and eigenanalysis, and a full three-dimensional Cosserat rod simulation for capturing large deformations and geometric nonlinearities.

\subsection{Deterministic IEC Beam Model}

To explore the mode selection mechanism (Eq.~\ref{eq:mode_selection}), we discretize the linearized beam equations using a finite difference scheme on a 1D domain $s \in [0, L]$. The rod is divided into $N=100$ segments. The information field $I(s)$ is mapped to local stiffness $E_i$ and rest curvature $\kappa_{i}$ at each node. We solve the resulting boundary value problem (BVP) using a standard shooting method (or sparse matrix solver for the eigenproblem). This allows rapid exploration of the $(\chi_\kappa, g)$ parameter space to identify regions where S-modes become the ground state.

\subsection{3D Cosserat Rod Implementation (PyElastica)}

For full 3D simulations, we utilize \texttt{PyElastica}~\cite{pyelastica_zenodo,gazzola2018forward}, an open-source Python implementation of Cosserat rod theory. Model parameters are listed in Table~\ref{tab:parameters}. The spine is modeled as a Cosserat rod with the following specifications:

\begin{itemize}
    \item \textbf{Discretization}: The rod is discretized into $n=50$--$100$ elements.
    \item \textbf{Information Field}: For spinal simulations, $I(s)$ is specified as a bimodal Gaussian distribution representing HOX-patterned identity:
    \begin{equation}
    I(s) = A_c \exp\left[-\frac{(s/L - s_c)^2}{2\sigma_c^2}\right] + A_l \exp\left[-\frac{(s/L - s_l)^2}{2\sigma_l^2}\right] + I_0,
    \label{eq:info_field_spinal}
    \end{equation}
    where $A_c = 0.5$, $s_c = 0.80$, $\sigma_c = 0.08$ (cervical lordosis), $A_l = 0.7$, $s_l = 0.25$, $\sigma_l = 0.10$ (lumbar lordosis), and $I_0 = 0.3$ (baseline). For scoliosis perturbations, a lateral asymmetry is added: $I(s) \to I(s) + \varepsilon_{\mathrm{asym}} \exp[-(s/L - 0.6)^2/(2 \cdot 0.08^2)]$ with $\varepsilon_{\mathrm{asym}} = 0.01$--$0.05$.
    \item \textbf{IEC Coupling}: We implemented a custom callback in \texttt{PyElastica} that updates the local rest curvature vector $\bm{\kappa}^0(s)$ and bending stiffness matrix $\mathbf{B}(s)$ at each time step (or initialization) based on the information field $I(s)$.
    \item \textbf{Boundary Conditions}: The rod is clamped at the base (sacrum) and free at the top (cranium), simulating a cantilever column under gravity. For specific validation cases, clamped-clamped conditions are used.
    \item \textbf{Gravitational Loading}: Gravity is applied as a uniform body force $\mathbf{f} = \rho A \mathbf{g}$.
    \item \textbf{Damping}: To find static equilibrium configurations, we apply external damping ($\nu \sim 0.1$--$1.0$ s$^{-1}$) and integrate the dynamic equations until the kinetic energy dissipates ($v_{\max} < 10^{-6}$ m/s).
\end{itemize}

The source code for the IEC-modified Cosserat solver is available in the \texttt{spinalmodes} Python package (see Data Availability).

\subsection{Parameter Sweeps and Mode Classification}

We perform systematic parameter sweeps over the coupling strength $\chi_\kappa$ (range $[0, 0.1]$) and gravitational acceleration $g$ (range $[0.01, 1.0]$ $g_{\mathrm{Earth}}$). For each simulation, we compute the equilibrium shape and evaluate the following metrics:
1.  \textbf{Geodesic Deviation} $\widehat{D}_{\mathrm{geo}}$: Quantifies the difference between the realized shape and the gravity-only geodesic.
2.  \textbf{Lateral Deviation} $S_{\mathrm{lat}}$: Measures symmetry breaking in the coronal plane.
3.  \textbf{Cobb Angle}: Standard clinical measure for scoliotic curves.

Regimes are classified based on the normalized geodesic deviation $\widehat{D}_{\mathrm{geo}}$. We define the \emph{gravity-dominated} regime ($\widehat{D}_{\mathrm{geo}} < 0.1$) as the region where gravitational potential energy dominates, leading to monotonic sag. The \emph{cooperative} regime ($0.1 \leq \widehat{D}_{\mathrm{geo}} \leq 0.3$) corresponds to the range where information and gravity balance to produce a stable, counter-curved S-shape. The \emph{information-dominated} regime ($\widehat{D}_{\mathrm{geo}} > 0.3$) is reached when information-driven metric distortions become the primary determinant of geometry, which, as we show, also coincides with the onset of symmetry-breaking instabilities (scoliosis-like modes). These thresholds were determined by analyzing the bifurcation of the lowest eigenvalue $\lambda_0$ and the emergence of non-monotonic curvature profiles.

\subsection{Multi-Scale Protein Mechanics Pipeline}

To ground the macroscopic parameters $\eta_p, \eta_a, \Gamma_m$ in molecular reality, we developed a multi-scale pipeline integrating AlphaFold predictions with structural energetics and tissue-level homogenization.
\begin{enumerate}
    \item \textbf{Structure Prediction:} We obtained AlphaFold-predicted structures for 23 proteins implicated in spinal mechanotransduction and energy metabolism (see Supplementary Table S2 for UniProt IDs and pLDDT scores). Structural anisotropy was quantified as the ratio of principal moments of inertia of the C$\alpha$ backbone, a computationally tractable proxy for conformational asymmetry and cytoskeletal maintenance cost. All structures with pLDDT $< 50$ at critical functional domains were excluded.
    \item \textbf{Molecular Mechanics (Steered MD):} We subjected the AlphaFold structures to simulated mechanical stress using Steered Molecular Dynamics (SMD) in GROMACS (version 2023.1, CHARMM36m force field, TIP3P water model). A constant pulling velocity ($0.01$ nm/ps) and harmonic spring constant ($1000$ kJ/mol/nm$^2$) were applied to terminal residues. Single-molecule stiffness $k_{\mathrm{molecule}}$ was extracted from the linear regime of the force-extension curves.
    \item \textbf{Fibril Assembly Coarse-Graining:} Single-molecule properties and RNA-seq-derived stoichiometries were coarse-grained into fibril assemblies. For collagen, we applied standard D-band periodicity packing geometry to calculate an effective fibril modulus $E_{\mathrm{fibril}}$.
    \item \textbf{Tissue Homogenization:} We employed mixture theory to compute the effective tissue stiffness tensor $\mathbf{C}_{\mathrm{tissue}} = \sum \phi_i \mathbf{E}_i$, where volume fractions $\phi_i$ were estimated from histological literature. Anisotropy was incorporated based on polarized light microscopy data of collagen orientation.
    \item \textbf{Validation:} Pipeline predictions were validated against atomic force microscopy (AFM) nanoindentation of single fibrils and bulk mechanical testing of vertebral tissue, yielding strong agreement ($R^2 = 0.81$).
\end{enumerate}

\subsection{Tensor Elasticity and Vector Field Analysis}
To capture the directional instabilities inherent in Information-Elasticity Coupling, we developed a tensor mechanics framework. The isotropic stiffness $E(s)$ was replaced with an orientation-dependent effective stiffness:
\begin{equation}
E_{\mathrm{eff}}(s, \theta) = E_{\parallel} \cos^2(\theta) + E_{\perp} \sin^2(\theta),
\end{equation}
where $\theta$ is the angle between the information gradient direction and the principal stress direction. We calculated an alignment parameter $\alpha(s) = |\hat{n}_{\mathrm{info}}(s) \cdot \hat{n}_{\mathrm{stress}}(s)|$, measuring tissue anisotropy via polarized imaging and computational vector field mapping of HOX expression. This permitted the identification of vector mismatch regions where $\alpha(s) \to 0$, identifying domains of scoliotic vulnerability.

\subsection{Protein Dynamics and Energy Modeling}

We simulated the temporal competition between protein classes using a system of coupled differential equations (Eq.~\ref{eq:protein_dynamics}). The model tracks the energy pool $E(t)$ available for protein synthesis/maintenance over the 5-20 year age range.
\begin{itemize}
    \item \textbf{Supply:} Modeled as $S(t) = S_0 (L(t)/L_0)^2$, reflecting the diffusion-limited transport through the vertebral endplate surface area.
    \item \textbf{Demand:} Modeled as $D(t) = D_{maint} (L/L_0) + D_{grav} (L/L_0)^4$, capturing the linear scaling of tissue volume maintenance and the quartic scaling of gravitational moment opposition.
    \item \textbf{Dynamics:} The fidelity of the mechanosensory array $F(t)$ evolves according to $dF/dt = k_{repair}(S - D)$ when $S > D$, and $dF/dt = k_{damage}(S - D)$ when $S < D$ (deficit), bounded to $[0, 1]$.
\end{itemize}

\subsection{Numerical convergence and validation}

A convergence study was performed by varying the number of elements $n$ from $10$ to $200$. We found that the normalized geodesic deviation $\widehat{D}_{\mathrm{geo}}$ converges to within 1\% for $n \geq 50$ (see Supplementary Fig.~S1). The numerical implementation was validated against analytical solutions for small-deflection Euler-Bernoulli beams. The \texttt{PyElastica} implementation was further verified by reproducing standard buckling and hanging chain benchmarks, with error metrics $||\mathbf{r}_{num} - \mathbf{r}_{ana}||/L < 10^{-4}$.
