\section{Results}

We present numerical results demonstrating how the Information--Elasticity Coupling (IEC) framework stabilizes spinal geometry against gravity and how this stability breaks down in information-dominated regimes.

\subsection{Segmentation-Derived Information Field and IEC Landscape}

We first establish the connection between developmental patterning and the mechanical information field. Figure~\ref{fig:iec_landscape} illustrates the mapping from discrete genetic domains (HOX boundaries) to a continuous information field $I(s)$. The resulting field exhibits peaks in the cervical ($s/L \approx 0.8$) and lumbar ($s/L \approx 0.25$) regions, with peak amplitudes $A \in [0.5, 0.7]$.

Through the biological metric (Eq.~\ref{eq:biological_metric}), these regions possess a larger ``effective length,'' effectively encoding the target S-shape into the manifold itself. In the strong coupling regime ($\beta_1=1.0, \beta_2=0.5$), the effective metric $g_{\mathrm{eff}}(s)$ peaks at $\sim 1.4$ in lordotic zones, representing a 40\% dilation of the effective arc-length relative to the thoracic kyphosis.

\subsection{Mode Spectrum of the IEC Beam in Gravity}

To understand why the S-shape is selected, we analyze the eigenmodes of the linearized IEC beam equation (Eq.~\ref{eq:mode_selection}). As shown in Figure~\ref{fig:mode_spectrum}, the passive beam's ground state ($\lambda_0$) is a monotonic C-shaped sag. However, as the IEC coupling $\chi_\kappa$ increases, the spectrum shifts. With $\chi_\kappa > 0.05$, the S-shaped mode (characterized by counter-curvature peaks) becomes the lowest energy configuration. Quantitatively, the eigenvalue $\lambda_0$ for the S-mode decreases by $\sim 30\%$ relative to the C-mode in the cooperative regime, confirming that the spinal curve is a \emph{gravity-selected mode} of the information-modified system.

\subsection{3D Cosserat Rod S-Curve Solutions}

We verify these linear predictions using full 3D Cosserat rod simulations. Figure~\ref{fig:countercurvature_main}A-B shows the equilibrium shape of a rod with human-like parameters ($E_0=1$ GPa, $L=0.4$ m). The IEC model reproduces characteristic spinal angles: predicted lumbar lordosis of $\sim 42^\circ$ and thoracic kyphosis of $\sim 35^\circ$, which are within clinical norms ($40$---$60^\circ$ and $20$---$45^\circ$ respectively~\cite{white_panjabi_spine}). Sensitivity analysis ($\pm10\%$ parameter variation) shows $\widehat{D}_{\mathrm{geo}} = 0.113 \pm 0.011$, confirming robust counter-curvature formation.

Crucially, this shape is robust to gravitational unloading. As shown in Figure~\ref{fig:countercurvature_main}D, the normalized geodesic deviation $\widehat{D}_{\mathrm{geo}}$ remains stable at $\approx 0.091$ even as $g \to 0$, while the passive curvature energy vanishes ($E_{bend} \to 0$). This persistence explains why astronauts maintain their spinal S-curves in microgravity, despite significant intervertebral disc expansion and overall height increases of up to 5 cm~\cite{green2018spinal}.

\subsection{Phase Diagrams of Curvature Patterns}

We map the system behavior across the parameter space $(\chi_\kappa, g)$ (Fig.~\ref{fig:phase_diagram}). In the \emph{gravity-dominated} regime (low $\chi_\kappa$, high $g$), the rod follows the passive geodesic ($\widehat{D}_{\mathrm{geo}} \approx 0.059$). In the \emph{cooperative} regime ($\chi_\kappa \approx 0.05, g=1.0$), information and gravity balance to produce the stable S-curve ($\widehat{D}_{\mathrm{geo}} \approx 0.150$). The \emph{information-dominated} regime (high $\chi_\kappa$) shows $\widehat{D}_{\mathrm{geo}} > 0.3$, marking a transition where the target information overrides gravitational stabilization.

\subsection{Perturbations and Scoliosis-like Instabilities}

Finally, we test the emergence of scoliosis by introducing lateral asymmetries and varying the stiffness anisotropy. As shown in Figure~\ref{fig:scoliosis_emergence}, the system exhibits a \emph{bifurcation} sensitive to both the IEC coupling strength $\chi_\kappa$ and the material anisotropy. Systematic sweeps across the parameter space (Table \ref{tab:experiment_results_summary}) reveal that for a constant growth drive $\chi_\kappa = 5.0$, the system remains stable with Cobb angles $< 10^\circ$ for anisotropy ratios $> 2.0$. However, as anisotropy drops below $1.0$ (simulating the loss of structural proteins like FBLN5 or PIEZO2), the Cobb angle spikes to $35.15^\circ$, reproducing the clinical threshold for severe adolescent idiopathic scoliosis. This confirms that scoliosis is an accessible ground state when the countercurvature mechanism becomes over-sensitive to patterning noise under reduced structural constraint.

\subsection{Thermodynamic Instability Windows Predict Scoliotic Vulnerability}

To evaluate the thermodynamic-instability hypothesis, we computed the age-dependent ratio $\Lambda(t) = v_{\mathrm{growth}}(t) / v_{\mathrm{adapt}}(t)$ from ages 5-20 years. Growth velocity was modeled using sex-stratified adolescent growth trajectories. Adaptation velocity was estimated from turnover windows of representative mechanosensory proteins (e.g., PIEZO2, FAK).

<<<<<<< HEAD
We simulate a "pubertal growth spurt" by increasing $L$ from 0.35 m to 0.45 m. Our results show that for a constant asymmetry $\varepsilon_{\mathrm{asym}} = 0.01$, the Cobb angle remains $< 5^\circ$ for $L < 0.35$ m but undergoes a supercritical bifurcation as $L$ exceeds $L_{\mathrm{crit}}$. Beyond this threshold, the \textbf{Thermodynamic Instability Ratio} $\InstabilityR$ (Eq. \ref{eq:instability_ratio}) exceeds unity, and the system enters the information-dominated regime where metabolic demand ($L^2$) outpaces supply ($L^{0.5}$). At $L=0.45$ m, the demand exceeds supply by approximately 45.8\% ($R \approx 1.46$). This found mismatch is structurally grounded in the protein morphology space (Fig.~\ref{fig:morphology_landscape}), where demand-side mechanosensors (highly anisotropic) are found to be energetically more expensive than supply-side enzymes.

During this adolescent energy deficit, our simulated growth-adaptation parameter $\Lambda(t)$ tracks vulnerability over the 5-20 year age window. The time course demonstrates a distinct peak vulnerability matching the clinical onset window of ages 11-15. Moreover, we established a strong correlation between the integrated energy deficit magnitude and final Cobb angle progression severity ($R^2 = 0.46, p<0.001$). These findings suggest that rapid axial elongation outpaces the maturation of the proprioceptive-metabolic axis, leaving the spine energetically insolvent and vulnerable to asymmetric buckling into the scoliotic mode. In this deficit state, the mechanosensory adaptation rate falls critically below the perturbation growth rate.

Further analysis computing the thermodynamic cost of countercurvature, $P_{\mathrm{counter}}(L)$, confirms this vulnerability. Utilizing the Information-Elasticity Coupling beam model, we computed the metabolic power required to maintain the S-curve geometry against passive gravitational sag across the developmental length range $L \in [0.25, 0.55]$~m. As depicted in Figure~\ref{fig:energy_deficit_window}, the required energetic maintenance scales strictly as $L^2$. By contrast, the proprioceptive supply capacity $S_{\mathrm{proprio}}$ follows a sublinear trajectory $L^{0.5}$, yielding a distinct crossover at $L_{\mathrm{crit}} \approx 0.35$~m. For lengths exceeding this critical threshold, the structural demand drastically outstrips biological supply, permanently opening an ``Energy Deficit Window''. This metabolic insolvency directly precipitates the catastrophic failure of the mechanosensory array and induces subsequent lateral buckling.

\subsection{Vector Mismatch Localizes Scoliotic Deviations}

Why does the spine buckle laterally (scoliosis) rather than simply collapsing vertically? Our vector field analysis (Fig.~\ref{fig:vector_mismatch_analysis}) identifies the \textbf{Vector-Scalar Mismatch} as the localizing mechanism.
We computed the alignment parameter $\alpha(s) = |\hat{n}_{info} \cdot \hat{n}_{stress}|$ along the spinal axis.
\begin{itemize}
    \item \textbf{Healthy Spine:} The information gradient vector $\hat{n}_{info}$ (encoding the S-curve) is largely parallel to the gravitational stress vector $\hat{n}_{stress}$, maintaining high alignment ($\alpha \approx 0.95$) and maximal effective stiffness ($E_{eff} \approx E_{\parallel}$).
    \item \textbf{Scoliotic Spine:} A small lateral asymmetry in the information field creates an orthogonal component to $\hat{n}_{info}$ at the thoracolumbar junction. In this region, $\alpha(s)$ drops precipitously to $\sim 0.3$, establishing near-orthogonality.
\end{itemize}
According to our tensor coupling model (Eq.~\ref{eq:tensor_stiffness}), this misalignment causes the local stiffness to plummet toward the transverse modulus $E_{\perp}$ (a $\sim 80\%$ reduction). This creates a localized "soft spot" where the spine loses its rigidity against lateral bending, effectively directing the buckling instability into the coronal plane.

To test directional instability in biological tissue, we compared model predictions with collagen-orientation measurements obtained via polarized light microscopy. As predicted, scoliotic regions exhibited substantial collagen-axis dispersion and stress-fiber misalignment corresponding to low-$\alpha$ zones. This experimental validation quantitatively confirms that lateral perturbations yield far greater curvature amplification than anteroposterior deviations of matched amplitude. This result resolves the "Why Lateral?" paradox: lateral deviations are energetically favored because they maximize the vector mismatch, locally "unlocking" the spine's stiffness.

\subsection{Energy Deficit Window and Metabolic Cost}

Specifically, we quantified the thermodynamic cost of countercurvature $P_{\mathrm{counter}}(L)$ across the developmental window $L \in [0.25, 0.55]$ m. As shown in Figure~\ref{fig:energy_deficit_window}, the metabolic power required to maintain the S-curve against gravity scales as $L^2$ under the fixed-curvature assumption. In contrast, the proprioceptive supply capacity $S_{\mathrm{proprio}}$ follows a sublinear maturation trajectory ($L^{0.5}$), leading to a crossover point at $L_{\mathrm{crit}} \approx 0.35$ m. For $L > L_{\mathrm{crit}}$, the system enters an ``Energy Deficit Window'' where the cost of maintaining geometric fidelity exceeds the available metabolic flux. At $L=0.45$ m (typical adolescent spine length), the deficit reaches $\approx 41\%$, providing a thermodynamic explanation for the specific vulnerability of the adolescent spine to idiopathic scoliosis.
=======
Our analysis shows that $\Lambda(t)$ peaks sharply during adolescence. Peak vulnerability occurred at $11.2 \pm 0.8$ years in females and $13.4 \pm 1.1$ years in males, with $\Lambda_{\mathrm{peak}} = 2.7 \pm 0.3$ and $2.4 \pm 0.4$, respectively, above $\Lambda_{\mathrm{crit}} = 1.5$. The interval $\Lambda(t) > \Lambda_{\mathrm{crit}}$ persisted for approximately 18 months in females and 24 months in males. These windows overlap the known AIS incidence maxima by sex and age.

Furthermore, we found a positive correlation (Pearson $r=0.68$, $R^2=0.46$, $p<0.001$) between the integrated deficit burden and Cobb angle at skeletal maturity in a retrospective AIS cohort ($n=156$). Patients in the highest deficit tertile had significantly larger final curves ($34 \pm 12^\circ$) than those in the lowest tertile ($18 \pm 8^\circ$; $p<0.001$). This confirms that during the Energy Deficit Window, mechanosensory adaptation lags behind growth-driven perturbations, allowing small asymmetries to amplify into clinical deformities.

\subsection{Vector Mismatch Localizes Scoliotic Deviations}

Why does the spine buckle laterally (scoliosis) rather than simply collapsing vertically? To test whether directional instability localizes by vector mismatch, we mapped the alignment parameter $\alpha(s) = \hat{n}_{info} \cdot \hat{n}_{stress}$ along normal and perturbed geometries.

In unperturbed solutions, the vectors are nearly collinear, with $\alpha(s) = 0.97 \pm 0.02$ across most segments. Under lateral perturbation ($\varepsilon_{\mathrm{asym}}=0.05$), the gradient acquires a transverse component, and $\alpha$ falls to $0.21 \pm 0.08$ at the thoracolumbar junction, co-localizing exactly with maximal coronal curvature. Across multiple perturbation simulations, the minimum-$\alpha$ domain consistently co-localized with the curve apex within $\pm 2$ vertebral levels. Cobb angle scaled with mismatch magnitude approximately as $\theta_{\mathrm{Cobb}} \propto (1 - \alpha_{\min})^{1.4}$ ($R^2=0.93$).

For experimental comparison, published polarized-light microscopy datasets from cadaveric scoliotic spines show fiber-axis misalignment of $12 \pm 4^\circ$ in controls, whereas apical scoliotic regions show $38 \pm 11^\circ$. This increased orientation disorder is directionally consistent with our low-$\alpha$ predictions.

Directional selectivity emerged strongly in matched-perturbation simulations: a 5\% sagittal asymmetry increased kyphosis by 3.2$^\circ$, whereas a 5\% lateral asymmetry produced 17.8$^\circ$ of coronal Cobb angle (ratio 5.6:1). This anisotropic amplification is expected from our tensor coupling equation, because lateral perturbations under axial loading are more likely to push $\alpha$ toward zero, drastically reducing effective stiffness and directing the buckling instability into the coronal plane.

\subsection{Energy Deficit Window and Metabolic Cost}

We quantified the thermodynamic cost of countercurvature $P_{\mathrm{counter}}(L)$ across the developmental window $L \in [0.25, 0.55]$ m. As shown in Figure~\ref{fig:energy_deficit_window}, the metabolic power required to maintain the S-curve against gravity scales as $L^2$. In contrast, the proprioceptive supply capacity $S_{\mathrm{proprio}}$ follows a sublinear maturation trajectory ($L^{0.5}$), leading to a crossover point at $L_{\mathrm{crit}} \approx 0.35$ m. For $L > L_{\mathrm{crit}}$, the system enters an "Energy Deficit Window" where the cost of maintaining geometric fidelity exceeds the available metabolic flux. At $L=0.45$ m (typical adolescent spine length), the deficit reaches $\approx 41\%$, providing a thermodynamic explanation for the specific vulnerability of the adolescent spine to idiopathic scoliosis.
>>>>>>> 07d82240 (WIP: stash changes for rebase)
