\newpage
\appendix
\section*{Supplementary Information}

\setcounter{section}{0}
\renewcommand{\thesection}{S\arabic{section}}
\renewcommand{\theequation}{S\arabic{equation}}
\setcounter{equation}{0}
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{table}{0}

\section{Cosserat Rod Geometry and Kinematic Parameterization}

We model the human spine as a Cosserat rod, defined by a centerline curve $\mathbf{r}(s,t) \in \mathbb{R}^3$ parameterized by arc length $s \in [0, L]$. This dimensionality reduction from 3D elasticity to a 1D director theory is justified by the high aspect ratio of the vertebral column ($L/d \approx 13$). To each point $s$, we attach an orthonormal material frame $\{\mathbf{d}_1(s,t), \mathbf{d}_2(s,t), \mathbf{d}_3(s,t)\}$ describing the cross-sectional orientation~\cite{goriely2017mathematics, gazzola2018forward}. The kinematics are governed by:
\begin{align}
\mathbf{r}'(s) &= \mathbf{v}(s) = \mathbf{d}_3(s) + \boldsymbol{\varepsilon}(s), \\
\mathbf{d}_i'(s) &= \mathbf{u}(s) \times \mathbf{d}_i(s),
\end{align}
where $\mathbf{v}(s)$ is the tangent vector, $\boldsymbol{\varepsilon}(s)$ represents shear and extension strains, and $\mathbf{u}(s) = \kappa_1 \mathbf{d}_1 + \kappa_2 \mathbf{d}_2 + \tau \mathbf{d}_3$ is the Darboux vector encoding flexural curvature $\kappa_{1,2}$ and torsional twist $\tau$.

\paragraph{Boundary Conditions.} The sacral base ($s=0$) is clamped and the cranial tip ($s=L$) is free:
\begin{align}
\mathbf{r}(0) = \mathbf{0}, \quad &\mathbf{d}_i(0) = \mathbf{e}_i \quad (\text{Clamped Base}) \nonumber \\
\mathbf{n}(L) = \mathbf{0}, \quad &\mathbf{m}(L) = \mathbf{0} \quad (\text{Free Tip})
\end{align}

\section{Spinal Modal Analysis and the Critical Delay Parameter ($\tau_c$)}

The linear stability of the active spine is evaluated using the Delay Differential Equation (DDE) described in the main text. Substituting the separable solution $\kappa(s,t) = y(s) e^{\lambda t}$ into the linearized DDE yields the spatial boundary value problem for a clamped-free rod:
\begin{equation}
EI y''''(s) = - (\rho A \lambda^2 + \eta \lambda + (K_P + K_D \lambda) e^{-\lambda \tau}) y(s) \equiv \mu y(s)
\end{equation}

The non-trivial solutions correspond to characteristic roots $\beta_n$ satisfying $\cosh(\beta_n L) \cos(\beta_n L) + 1 = 0$, giving $\beta_1 L \approx 1.875$, $\beta_2 L \approx 4.694$, and $\beta_3 L \approx 7.855$. Evaluating the clamped-free spatial eigenvectors $y_n(s)$, the dominant internal elastic mode (Mode 2) exhibits maximum localized curvature at $s/L \approx 0.53$ and maximum transverse displacement at $s/L \approx 0.47$. Mode 3 exhibits curvature peaks at $s/L \approx 0.31$ and $0.71$. These bounds accurately capture the standard thoracic apex of right-convex Adolescent Idiopathic Scoliosis (AIS) curves, validating the correspondence between the spatial buckling modes and clinical presentations.

By isolating the temporal spectrum $\mu = EI \beta_n^4$, the stability boundary along the imaginary axis $\lambda = i\omega$ satisfies:
\begin{align}
\tan(\omega \tau_c) &= \frac{K_D \omega}{K_P}, \\
\rho A \omega^2 - EI \beta_n^4 &= K_P \cos(\omega \tau_c) + K_D \omega \sin(\omega \tau_c).
\end{align}

Given physical trunk parameters and generalized stretch reflex delay ($40 \sim 90$ ms in human adolescents), stability analysis of this inverted pendulum analog typically localizes the critical delay $\tau_c$ near $50-100$ ms. During the adolescent growth spurt, height velocity outpaces axonal myelination and expansion, pushing absolute neural delay $\tau$ across this $\tau_c$ threshold and inducing a supercritical Hopf bifurcation (the Derivative Gain Trap).

\section{Cosserat Force and Moment Balance}

For a static rod subject to gravity $\mathbf{f}_g = \rho A \mathbf{g}$ and active moments induced by the information field, the balance equations for internal force $\mathbf{n}$ and moment $\mathbf{m}$ are:
\begin{align}
\mathbf{n}'(s) + \mathbf{f}_g &= \mathbf{0}, \nonumber \\
\mathbf{m}'(s) + \mathbf{r}'(s) \times \mathbf{n}(s) + \mathbf{m}_{\mathrm{info}}'(s) &= \mathbf{0},
\end{align}
where $\mathbf{m}_{\mathrm{info}}$ represents the active couple generated by the information field gradient.

\section{Developmental Information Field Parameterization}

The scalar information field $I(s)$ is parameterized as a bimodal Gaussian:
\begin{equation}
I(s) = A_c \exp\left[-\frac{(s/L - s_c)^2}{2\sigma_c^2}\right] + A_l \exp\left[-\frac{(s/L - s_l)^2}{2\sigma_l^2}\right] + I_0,
\end{equation}
where $A_c = 0.5$, $s_c = 0.80$, $\sigma_c = 0.08$ (cervical lordosis), $A_l = 0.7$, $s_l = 0.25$, $\sigma_l = 0.10$ (lumbar lordosis), and $I_0 = 0.3$ (baseline).

The information content is quantified using the local Shannon entropy:
\begin{equation}
\ShannonH[I] = -\int I(s) \log I(s) \, ds,
\end{equation}
which bounds the precision of the target metric specification.

\section{Biological Metric and Information Geometry}

The central hypothesis of the IEC framework is that developmental information modifies the effective metric experienced by the rod's reference manifold. We define a biological metric:
\begin{equation}
d\ell_{\mathrm{eff}}^2 = g_{\mathrm{eff}}(s)\,ds^2 = \exp\left[2\left(\beta_1 \tilde{I}(s) + \beta_2 \frac{\partial \tilde{I}}{\partial s}\right)\right] ds^2,
\end{equation}
where $\tilde{I}$ is the normalized information field and $\beta_{1,2}$ are dimensionless coupling constants.

\section{Geodesic Deviation and Regime Classification}

We quantify the degree of biological counter-curvature using the Kullback--Leibler divergence between the IEC-coupled geometry and the passive gravitational state:
\begin{equation}
\KLDivergence = \int_0^L \left[ \frac{1}{2} \left( \kappa_{\mathrm{IEC}}(s) - \kappa_{\mathrm{passive}}(s) \right)^2 \FisherInfo(s) \right] w(s)\, ds,
\end{equation}
where $\FisherInfo(s)$ is the local Fisher Information and $w(s) = g_{\mathrm{eff}}(s)$. The normalized geodesic deviation is:
\begin{equation}
\Dgeohat = \frac{\KLDivergence}{\KLDivergence^{\mathrm{max}}(\chi_\kappa, g)}.
\end{equation}

Regime thresholds: gravity-dominated ($\Dgeohat < 0.1$), cooperative ($0.1 \leq \Dgeohat \leq 0.3$), information-dominated ($\Dgeohat > 0.3$). These were determined by analyzing the bifurcation of the lowest eigenvalue $\lambda_0$.

\section{Metabolic Power Density and Dissipation}

The metabolic power density required to sustain counter-curvature is:
\begin{equation}
\dot{\mathcal{F}} = \int_0^L \left[ \eta_p \left|\frac{\partial \kappa}{\partial t}\right|^2 + \eta_a \left(\kappa(s) - \kappa_{\mathrm{passive}}(s)\right)^2 + \Gamma_m(s) \right] ds,
\end{equation}
where $\eta_p$ is the proprioceptive feedback dissipation coefficient, $\eta_a$ is the active moment maintenance coefficient, and $\Gamma_m(s)$ is the basal tissue maintenance cost density.

\section{Protein Dynamics and Energy Competition}

During adolescent growth, mechanosensory and growth/maintenance protein classes compete for a finite metabolic energy pool:
\begin{equation}
E_{total}(t) = E_{mechano}(t) + E_{growth}(t) + E_{metabolic}
\end{equation}
\begin{equation}
\frac{dE_{mechano}}{dt} = k_{syn,m} \cdot [Resources] - k_{deg,m} \cdot [Mechano] - \dot{\mathcal{F}}_{load}
\end{equation}
where $\dot{\mathcal{F}}_{load} \propto L^4$ represents the increasing thermodynamic load. The growth protein synthesis rate is driven by GH/IGF-1 and scales with $L^2$ (surface area of nutrient diffusion), creating the scaling mismatch at the molecular level.

The vulnerability is quantified via:
\begin{equation}
\Lambda(t) = \frac{dV/dt}{\tau_{adapt}}
\end{equation}
where $dV/dt$ is volumetric growth rate and $\tau_{adapt}$ is the mechanosensory adaptation time constant.

\section{The Spinal Clock and Convective Shutdown}

The intervertebral disc possesses an autonomous circadian clock driven by the BMAL1/CLOCK transcriptional loop~\cite{dudek2017clock}. Gravity provides the synchronization signal through daily loading cycles: daytime compression forces fluid out of the nucleus pulposus, while nocturnal unloading draws fluid back via osmotic pressure~\cite{urban2004nutrition}.

Under microgravity or prolonged bed rest, this pump arrests (``Convective Shutdown''), creating hypoxia that stabilizes HIF-1$\alpha$ and upregulates catabolic enzymes (MMP1, MMP3). The resulting matrix degradation reduces $B_{\mathrm{eff}}(s)$ and widens the Energy Deficit Window.

This mechanism generates three falsifiable predictions:
\begin{enumerate}
    \item Clock gene desynchronization (Per2, Bmal1 flattening in IVD) within 72 hours of unloading, preceding structural changes.
    \item HIF-1$\alpha$ stabilization in the nucleus pulposus preceding scoliotic curvature; HIF-1$\alpha$ inhibitors should delay onset.
    \item Pulsatile axial compression at the appropriate circadian phase will preserve spinal geometry more effectively than static compression.
\end{enumerate}
